### 2024-06-18 basic-statistics.qmd
20240618 - edits

In science, using what is called null-hypothesis significance testing (NHST), we typically consider an effect to be *statistically significant* if the *p*-value is less than .05.
That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect at least as extreme as the effect observed, if the null hypothesis were true.
So, you might expect around 5% of tests where the null hypothesis is true to be significant just by chance.

If the *p*-value is less than .05, we reject the null hypothesis ($H_0$) that there was no difference or association.
Thus, we conclude that there was a significant (non-zero) difference or association.
If the *p*-value is greater than .05, we fail to reject the null hypothesis; the difference/association was not significant.
Thus, we do not have confidence that there was a difference or association.
However, we do not accept the null hypothesis; it could be there we did not observe an effect because we did not have adequate power to detect the effect—e.g., if the effect size was small, the data were noisy, and the sample size was small and/or unrepresentative.

@@

Twelve Misconceptions About *p*-Values from @Goodman2008. Goodman also provides a discussion about why each is false. {#tbl-pValueMisconceptions}

>>>>

In science, using what is called null-hypothesis significance testing (NHST), we typically consider an effect to be *statistically significant* if the *p*-value is less than .05.
That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect at least as extreme as the effect observed, if the null hypothesis were true.
So, you might expect around 5% of tests where the null hypothesis is true to be statistically significant just by chance.

If the *p*-value is less than .05, we reject the null hypothesis ($H_0$) that there was no difference or association.
Thus, we conclude that there was a statistically significant (non-zero) difference or association.
If the *p*-value is greater than .05, we fail to reject the null hypothesis; the difference/association was not statistically significant.
Thus, we do not have confidence that there was a difference or association.
However, we do not accept the null hypothesis; it could be there we did not observe an effect because we did not have adequate power to detect the effect—e.g., if the effect size was small, the data were noisy, and the sample size was small and/or unrepresentative.

@@

: Twelve Misconceptions About *p*-Values from @Goodman2008. Goodman also provides a discussion about why each is false. {#tbl-pValueMisconceptions}
